### 2010-07-13 Winds/papers/Theory/First/wind_simple.tex
profx
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5038 ac11df5e-a788-c23b-a8e1-c758c323924d

\documentclass[12pt,preprint]{aastex}
\usepackage{natbib,amsmath}

\special{papersize=8.5in,11in}
\begin{document}

\newcommand{\taud}{$\tau_{\rm dust}$}
\newcommand{\mtaud}{\tau_{\rm dust}}
\newcommand{\maconfig}{2p$^6$3s}
\newcommand{\mbconfig}{2p$^6$3p}
\newcommand{\aconfig}{a~$^6$D$^0$}
\newcommand{\zconfig}{z~$^6$D$^0$}
\newcommand{\mvr}{v_{\rm r}}
\newcommand{\naid}{\ion{Na}{1}~$\lambda\lambda 5891, 5897$}
\newcommand{\mgiid}{\ion{Mg}{2}~$\lambda\lambda 2796, 2803$}
\newcommand{\mgiia}{\ion{Mg}{2}~$\lambda 2796$}
\newcommand{\mgiib}{\ion{Mg}{2}~$\lambda 2803$}
\newcommand{\feiia}{\ion{Fe}{2}~$\lambda 2586$}
\newcommand{\feiib}{\ion{Fe}{2}~$\lambda 2600$}
\newcommand{\feiic}{\ion{Fe}{2}$^* \; \lambda 2612$}
\newcommand{\feiie}{\ion{Fe}{2}$^* \; \lambda 2626$}
\newcommand{\feiid}{\ion{Fe}{2}~$\lambda\lambda 2586, 2600$}
\newcommand{\nmg}{$n_{\rm Mg^+}$}
\newcommand{\mnmg}{n_{\rm Mg^+}}
\newcommand{\nfe}{$n_{\rm Fe^+}$}
\newcommand{\mnfe}{n_{\rm Fe^+}}
\def\hub{h_{72}^{-1}}
\def\umfp{{\hub \, \rm Mpc}}
\def\mzq{z_q}
\def\zabs{$z_{\rm abs}$}
\def\mzabs{z_{\rm abs}}
\def\intl{\int\limits}
\def\cmma{\;\;\; ,}
\def\perd{\;\;\; .}
\def\ltk{\left [ \,}
\def\ltp{\left ( \,}
\def\ltb{\left \{ \,}
\def\rtk{\, \right  ] }
\def\rtp{\, \right  ) }
\def\rtb{\, \right \} }
\def\sci#1{{\; \times \; 10^{#1}}}
\def \rAA {\rm \AA}
\def \zem {$z_{\rm em}$}
\def \mzem {z_{\rm em}}
\def\smm{\sum\limits}
\def \cmm  {cm$^{-2}$}
\def \cmmm {cm$^{-3}$}
\def \kms  {km~s$^{-1}$}
\def \mkms  {{\rm km~s^{-1}}}
\def \lyaf {Ly$\alpha$ forest}
\def \Lya  {Ly$\alpha$}
\def \lya  {Ly$\alpha$}
\def \mlya  {Ly\alpha}
\def \Lyb  {Ly$\beta$}
\def \lyb  {Ly$\beta$}
\def \lyg  {Ly$\gamma$}
\def \ly5  {Ly-5}
\def \ly6  {Ly-6}
\def \ly7  {Ly-7}
\def \nhi  {$N_{\rm HI}$}
\def \mnhi  {N_{\rm HI}}
\def \lnhi {$\log N_{HI}$}
\def \mlnhi {\log N_{HI}}
\def \etal {\textit{et al.}}
\def \lyaf {Lyman--$\alpha$ forest}
\def \mnmin {\mnhi^{\rm min}}
\def \nmin {$\mnhi^{\rm min}$}
\def \O {${\mathcal O}(N,X)$}
\newcommand{\cm}[1]{\, {\rm cm^{#1}}}
\def \snrlim {SNR$_{lim}$}

\title{Simple Wind Models}

\author{
J. Xavier Prochaska\altaffilmark{1}, 
Others
%John M. O'Meara\altaffilmark{2}, 
%Gabor Worseck\altaffilmark{1} 
%\& Scott Burles\altaffilmark{3}
}
\altaffiltext{1}{Department of Astronomy and Astrophysics, UCO/Lick Observatory, University of California, 1156 High Street, Santa Cruz, CA 95064}
%\altaffiltext{2}{Department of Chemistry and Physics, Saint Michael's College.
%One Winooski Park, Colchester, VT 05439}

\begin{abstract}
\begin{itemize}
\item Emission is a generic feature of (nearly) isotropic winds.  This
  emission fills in absorption at $v \approx 0$, significantly
  complicating the absorption-line analysis, especially of an ISM
  component. 
\item The relative strengths of the emission
lines, therefore, is sensitive to both the opacity and velocity
extent of the wind.
\end{itemize}
\end{abstract}

\keywords{absorption lines -- intergalactic medium -- Lyman limit systems -- SDSS}

\section{Introduction}

Nearly all gaseous objects in astrophysics that emit light are also
observed to drive gas away.  [word] This includes the jets of protostars, the
strong stellar winds of massive O and B stars, the gentle Solar wind
of our Sun, the associated absorption of bright quasars, and the
spectacular jets of radio-loud AGN.   These gaseous outflows moderate
the accretion of fresh material onto the object, inject energy and
momentum into gas on large scales, and ...
[Last line]

Star-burst galaxies, whose light is dominated by \ion{H}{2} regions
and massive stars, are also observed to drive gaseous outflows.  The
flows are generally expected (and sometimes observed) to have multiple
phases, with a hot and diffuse phase traced by X-ray emission
\citep[e.g.][]{strickland,martin+k} and a cool, denser phase traced by
H$\alpha$ emission \citep[e.g.][]{NGC891,M87}.

Beyond the local universe, galactic outflows are predominatly
revealted by UV and optical absorption lines, e.g.\ \ion{Na}{1},
\ion{Mg}{2}, \ion{Si}{2} and \ion{C}{4} transitions.  With the galaxy
as a backlight, one almost universally observes predominantly
blue-shifted material indicating a galactic-scale flow of gas toward
Earth and away from the galaxy.  These transitions are sensitive to
the cool (\ion{Mg}{2}; $T \sim 10^4$K) and warm (\ion{C}{4}; $T \sim
10^5$K) phases of the flow.  The incidence of cool gas outflows is
nearly universal in star-forming galaxies;  this includes systems at $z \sim 0$
which exhibit \ion{Na}{1} and \ion{Mg}{2} absorption
\citep{rupke,martin,student}, $z \sim 1$ star-forming galaxies traced by
\ion{Fe}{2} and \ion{Mg}{2} transitions \citep{wcp+09,rubin+10}, and
$z>2$ Lyman break galaxies (LBGs) that show blue-shifted \ion{Si}{2},
\ion{C}{2}, and \ion{O}{1} transitions \citep{lowenthal,steidel94}.

Although the obsevation of metal-line absorption is a well established
means of identifying outflows at all epochs, little effort has been
made to compare these observations against (even idealized) wind
models.  The focus to date has been to gelan the limited information
affored by direct analysis of the absorption lines.  The data do yield
robust measurements of the speed of the gas yet poorly constrain the
optical depth, covering fraction, density, temperature, and distance
of the gas from the emitting regions.  Constraints related to the
mass, energetics, and momentum of the flow suffer from
orders of magnitude uncertainty.  Both the origin and impact of the
winds, therefore, remain open points of debate.

Recent studes of $z \sim 1$ star-forming galaixes have revealed that
the cool gas frequently exhibits significant \ion{Mg}{2} emission in
addition to the ubiquitous blue-shifted absorption
\citep{wcp+09,rubin+10b}.  \cite{rubin+10a} have also detected line
emission from non-resonance \ion{Fe}{2}~$^*$ transitions and spatially
extended \ion{Mg}{2} emission which they interpret as originating from
photons scattered off a galactic-scale wind.  They estimate that the
wind extends to at least 5\,kpc from the galaxy.  Line emission is
also observed for $z \sim 3$ LBGs via the resonant \lya\ transition
and fine-strcuture \ion{Si}{2}$^*$ transitions \citep{cb58,shapley}.
\cite{rubin+10a} emphasize that a comprehensive analysis of these data
(e.g.\ via integral-field-unit observations) could yield unique
diagnostics on the morphology and density of the outflow, eventually
yielding tigheter constrains on the energetics of the flow.  
The crucial piece missing from this path is a radiative transfer
treatment that properly ...

In this paper, we take the first steps toward modeling the absorption
and emission properties of cool gas outflows.  Using Monte Carlo
radiative transfer technqiues, we study the nature of \ion{Mg}{2} and
\ion{Fe}{2} absorption and emission for winds with a range of
properties.  Although the winds are idealized, the results frequently
contradict our initial intuition and caution one [word] on the
challenges of converting the observables to (even crude) physical
constraints.  [Add another line or two]

[Outline of paper]

[Point out IR emission from winds locally implies dust]

\section{Methodology}
\label{sec:method}

This section describes the methodology [employed] to generate
emission/absorption profiles from simple wind models.

\subsection{The Radiative Transitions}

In this paper, we focus on two sets of radiative transitions
arising from Fe$^+$ and Mg$^+$ ions
(Table~\ref{tab:atomic},Figure~\ref{fig:energy}).
Although this is a necessarily limited
set, the two ions and their transitions are characteristic
of the majority of low-ions 
observed in cool-gas outflows. Therefore, many
of the results may be generalized to observational studies that
consider other atoms and ions of cool gas.

The Mg$^+$ ion, with a single 3s electron in the ground-state,
exhibits an alkali doublet of transitions at $\lambda \approx
2800$\AA\ similar to the
\lya\ transitions of neutral hydrogen.  Figure~\ref{fig:energy}
presents the energy level diagram for this 
\mgiid\ doublet.  In non-relativistic quantum
mechanics, the 2p$^6$3p energy level is said to be split by spin-orbit
coupling which yields the observed line doublet.  These are the only
\ion{Mg}{2} electric-dipole transitions 
with wavelengths near 2800\AA\ and the transtion connecting
the $\rm {}^2P_{3/2}$ and $\rm {}^2P_{1/2}$ states is forbidden by several
selection rules.  Therefore, an absorption from
\maconfig~$\to$~\mbconfig\
is followed essentially 100$\%$ of the time by a spontaneous decay
($t_{\rm decay} \approx 4\sci{-9}$s to the
ground state. Our treatment will ignore any other possibilities
(e.g.\ absorption by a second photon when the electron is at the \mbconfig\ level).

The physics of the \mgiid\ doublet
is analagous to the physics of \ion{H}{1}
\lya, the \naid\ doublet, and many other doublets commonly
studied in the interstellar medium (ISM) of distant galaxies.  The
doublets differ only in 
their rest wavelengths and the energy of the doublet separation. 
For \ion{H}{1} \lya, the
separation is sufficiently small ($\Delta v = c \Delta E / E \approx
1.3 \, \mkms$) that most radiative transfer treatments actually ignore that
there are two transitions.   This is generally justifiable for \lya\ because 
most astrophysical processes have turbulent motions that
significantly exceed its velocity separation and effectively mix the
two transitions.  For \ion{Mg}{2} ($\Delta v \approx 770 \, \mkms$),  
\ion{Na}{1} ($\Delta v \approx 304 \, \mkms$), and most of the other doublets
commonly observed, the separation is large and the transitions
must be treated separately.  Regarding the analysis of outflows, the line
separation plays an important role when the outflow speed matches or
exceeds $\Delta v$.  [Explain?]

Iron exhibits the most complex set for energy levels of the elements
frequently studied in astrophysics.  The Fe$^+$ ion alone has over XXX
energy levels recorded \citep{iron}, and even this is an
incomplete list.  
One reason for iron's complexity is
that the majority of its configurations exhibit fine-structure splitting.
This includes the ground-state configuration (\aconfig) which is split
into 5 levels with 
excitation energies $T_{\rm ex} \equiv \Delta E / k$ ranging from
$T_{\rm ex} \approx 500-1500$\,K, labelled by the total angular momentum
$J$. The ground-state has $J=9/2$ (Figure~\ref{fig:energy}).  
Transitions between these fine-structure levels are 
forbidden (magnetic-dipole) with spontaneous decay times of several hours.

In this paper, we examine transitions between the ground-state
configuration and the energy levels of the \zconfig\
configuration.  This set of transitions (named the
UV1 mutliplet) have wavelengths near 2600\AA.
There are two resonance-line transitions\footnote{We adopt the
  standard convention that a ``resonance line'' is an electric-dipole
  transition connected to the ground-state.} 
associated with this multiplet (\feiid)
corresponding to $\Delta J = 0, -1$.  The solid (green) downward
arrows in Figure~\ref{fig:energy} mark the transitions that connect to
the upper energy levels of the resonance lines.  These transitons can
occur following the absorption of a single photon by Fe$^+$ in its
ground-state.

The Figure also shows (as dashed downward arrows) two of the
transitions that connect to higher energy levels of the \zconfig\
configuration.  Ignoring collisions and recombinations, these
transitions may only can only occur after the absorption
of two photons: one to raise the electron from the ground-state to an
excited state and one from the excited state to one of the \zconfig\
levels with $J \le 5/2$.  The excitation of these levels by
the absorption of UV photons is termed indirect UV pumping
\citep[e.g][]{sv02,pcb06} and requires the ion to lie
near an intense source of UV photons (see $\S$~\ref{sec:pump}).  In the
following, we will assume that this does not occur.

Our calculations also ignore collisional
processes\footnote{Recombination is also ignored.}, i.e.\ collisional
excitation and de-excitation of the various levels.  For the
fine-structure levels of the \aconfig\ configuration, the excitation
energies are modest ($T_{\rm ex} \sim 1000$\,K) but the critical
density $n_c$ is very large.  For the \aconfig$_{9/2} \to
$\aconfig$_{7/2}$ transition, the critical density $n_e^C \approx 4
\sci{5} \cm{-3}$.  At these densities, one would predict  
detectable quantities of \ion{Fe}{1} which has not yet been observed
in galactic-scale outflows
\citep[see also][]{pcb06}.  Furthermore, observations rarely show
{\it absorption} from the
fine-structure levels of the \aconfig\ configuration and that material
is not significantly blue-shifted \citep{rubin+10}.
Empirically, collisional excitation is insignificant and, by
inference, one may neglect collisional de-excitation.  
For Mg$^+$,$T_{\rm ex} \approx 50,000$\,K  
$n_e^C \approx 3 \sci{14} \cm{-3}$  [Maybe critical density is
not so interesting here...]
implying negligible collisional excitation for the cool gas.  
This implies the gas has zero opacity to the non-resonant lines.

\subsection{The Source}

Nearly all the absorption studies of cool-gas outflows 
have focused on intensely, star-forming galaxies.  The
intrinsic emisison of these galaxies is a complex combination of
stars and \ion{H}{2} regions that is then modulated by dust and gas
within the ISM.  For the spectral regions studied
here, most stars show a featureless continuum but a few spectral
types do show significant \ion{Mg}{2} and \ion{Fe}{2} absorption.
[Any emission??]  \ion{H}{2} regions, meanwhile, are observed to emit
at the \mgiid\ doublet, primarily due to recombinations in the outer
layers [How about FeII?].  It is beyond the scope of this paper to
properly model the 
stellar absorption and \ion{H}{2} region emission, but the reader
should be aware that they can complicate the observed spectrum,
independent of the galactic outflow, especially at velocities $v \approx 0 \mkms$.
In the following, we will simply assume a flat continuum 
normalized to unit value.  The size of the emitting
region $r_{\rm source}$ is a free parameter, although we tend to
restrict its value to be smaller than the minimum radial extent of any
gaseous component.

[What does MgI tell us about winds?  Why no P-Cygni?]

\subsection{Monte Carlo Algorithm}

[Describe 1D and 3D algorithms]

[Put tests in an Appendix?]

\subsection{Dust}
\label{sec:dust_method}

For the majority of models studied in this paper, we assume no dust.
This is an invalid assumption, espeically for material associated with
the ISM of the galaxy.  Because absorption line analysis is performed
on normalized spectra, the effects of dust are largely minimized.  For
scattered and resonantly trapped photons, however, the effects of dust
extinction can be significant.  Indeed, dust is frequently invoked to
explain the weak (or absent) \lya\ emission from star-forming galaxies
\citep[e.g.][]{shapley03}.  Although the transitions studied here have
significantly lower opacity than \lya, dust may play an important role
in the predicted profiles.

In a few models, we include the absorption by dust under the following
assumptions:
(i) the dust opacity scales with the density of the gas (i.e.\ we
adopt a fixed dust-to-gas ratio);
(ii) the opacity is independent of wavelength, a reasonable
approximation given the small spectral range analyzed;
(iii) the photons absorbed by dust are re-emitted at IR wavelengths
and `lost' from the system.  The dust absorption is normalized by \taud, 
the integrated opacity of dust from the center of the system to
infinity.  The ISM of a star-forming galaxy may be expected to exhibit
\taud\ of a few \citep[e.g.][]{steidel98}.

\section{The Fiducial Wind Model}
\label{sec:fiducial}

In this section, we study a simple yet illustrative wind model for
a galactic-scale outflow.  The properties of this wind were tuned, in
part, to yield a \ion{Mg}{2} absorption profile 
similar to that observed toward $z<1$, star-forming galaxies
\citep{wcp+09,rubin10b}.  We emphasize, however, that we do not
favor this fiducial model over any other wind scenario nor do its
properties have special physical motivation.
Its primary purpose is to establish a baseline
for discussion.

The fiducial wind follows a (mass-fluxing conserving) density law,

\begin{equation}
n_{\rm H} (r) = \frac{n^0_{\rm H}}{r^2} \;\;\; , 
\label{eqn:density}
\end{equation}
and a velocity

\begin{equation}
\vec v = v_r (r) \hat r = v_0 r \hat r \;\;\; , 
\label{eqn:vel}
\end{equation}
i.e., a purely radial flow.  Turbulent motions are
characterized by a Doppler parameter $b_{\rm turb}$.  
The wind is isotropic, dust-free, and extends from an inner wind
radius $r_{\rm inner}$ to an outer wind radius $r_{\rm outer}$.  
We scale the hydrogen density $n_{\rm H}$ to the number densities of
Mg$^+$ and Fe$^+$ ions by assuming solar relative abundances with an
absolute metallicity of 1/2 solar and depletion factors of 1/10 and
1/20 for Mg and Fe respectively, i.e.\  $\mnmg = 10^{-5.47} n_{\rm H}$ 
and \nfe=\nmg/2. At the center of the wind is a homogenous source of
continuum photons with a size $r_{\rm source}$. These parameters are
summarized in Table~\ref{tab:fiducial}.

In Figure~\ref{fig:fiducial_nvt} we plot the density and velocity
profiles against radius for the fiducial wind model;  
their single power-law expressions are evident.  The figure also
shows the optical depth profile for the \mgiia\ transition ($\tau_{2796}$).
Although plotted against radius, we first calculated $\tau_{2796}$ 
as a function of velocity by summing the opacity in a series of
discrete and small radial intervals
between $r_{\rm inner}$  and $r_{\rm outer}$.   We then mapped
$\tau_{2796}$ onto radius using the velocity law
(Equation~\ref{eqn:vel}). 
The $\tau_{2796}$ profile peaks at $\tau_{2796}^{\rm max} \approx 30$
with a velocity $v \approx 65 \mkms$ corresponding to $r \approx 1.3
r_{\rm inner}$.  The optical depth profile for the \mgiib\ transition
is scaled down by the $f\lambda$ ratio but otherwise identical.  Similarly,
the optical depth profiles for the \feiid\ transitions are
scaled down by $f \lambda$ and the \nfe/\nmg\ ratio.  
The adopted Doppler parameter
($b_{\rm turb} = 15 \mkms$) has only a modest impact on the results.
A larger/smaller $b$-value has minimal effect on the absorption
profiles and only tends to modify the widths and for resonant
transitions shifts the centroids of their emission lines.

Because the results are dictated by the $\tau_{2796}$ profile in
velocity space, the actual dimensions and density of the wind are
essentially unimportant provided they scale together to give nearly the same
$\tau_{2796}$ velocity profile.  Therefore, one may consider the choices for
$r_{\rm inner}, r_{\rm outer}, n_{\rm H}^0$ as arbitrary.
Nevertheless, we adopted values for this fiducial model with
some astrophysical motivation,  e.g. values that correspond to
galactic dimensions and a normalization that gives $\tau_{2796}^{\rm
  max} \sim 10$.

Using the methodology described in $\S$~\ref{sec:method}, we
propogated photons from the source and through the outflow to an
`observer' at $r \gg r_{\rm outer}$ that views the entire wind+source
complex.  Figure~\ref{fig:fiducial_1d} presents the 1D spectrum
that this observer would record, with the ``unattenuated'' flux
normalized to unit value.   The \ion{Mg}{2} doublet
shows the canonical `P-Cygni profile' that characterizes a continuum
source embedded within an outflow.  Strong absorption is evident at
$v < -50 \mkms$ in both transitions (equivalent widths $W_{2796} =
2.96$\AA\ and $W_{2803} = 1.29$\AA) and each shows emission at
positive velocities.  For this isotropic and dust-free model, the
total equivalent width of the doublet must be zero
(Table~\ref{tab:fiducial_EW}), i.e.\ every photon
absorbed eventually escapes the system, generally at lower
energies.  The wind simply shuffles the photons in frequency space.

Focusing on the \ion{Mg}{2} absorption, one notes the profiles lie
well above zero intensity and have similar depth even though their $f\lambda$
values differ by a factor of two.  In standard absorption line
analysis, this is generally 
the tell-tale signature of a `cloud' that has a high optical depth (i.e.\
saturated) which only partially covers the emitting source
\citep[e.g.][]{hamman10}.  Our fiducial wind model, however, 
{\it entirely covers the source}; the apparent partial covering must
be due to a different effect.
Figure~\ref{fig:noemiss} further emphasizes this point by comparing the 
absorption profiles from Figure~\ref{fig:fiducial_1d} against an
artificial model where no absorbed photons are 
re-emitted.   As expected from the
$\tau_{2796}$ profile (Figure~\ref{fig:fiducial_nvt}), the
`no-emission' model
produces a strong \mgiid\ doublet that absorbs all photons at
$v \approx -100 \mkms$, i.e.\ $I_{2796}^{\rm min} = \exp(-\tau^{\rm
  max}_{2796}) \approx 0$.
The fiducial model, meanwhile, has been `filled in' at $v \approx -100
\mkms$ by photons scattered and re-emitted by the wind (see below).  An
absorption-line analysis that ignores these effects
\citep[e.g.][]{sato} would (i) systematically underestimate the true optical
depth and (ii) falsely conclude that the wind partially covers the
source.  We will find that this
behaviour is a generic result of the wind models considered, even for cases that are
not fully isotropic.
[Dan's analytic calculation here? or Appendix?]

Turning to the emission profiles of the \mgiid\ doublet, one notes
that they are also and have comparable equivalent widths.  The
flux of the \mgiib\ transition even exceeds that for \mgiia\ giving a
line ratio that is far below the $2:1$ ratio that one may have naively
expected. 
[Does recombination predict 2:1??]
The emission profiles are very similar because the gas is optically
thick for a significant portion of the profile. 
The flux from \mgiib\ actually exceeds the
\mgiia\ emission because the wind speed is greater than the velocity separation
of the doublet, $|\mvr|_{\rm max} > (\Delta v)_{\rm MgII}$.
Therefore,
the \mgiib\ absorption profile partially absorbs the red wing of the
\mgiia\ emission profile.  The relative strengths of the emission
lines, therefore, is sensitive to both the opacity and velocity
extent of the wind.  We also emphasize that departures from a 2:1
ratio indicate that processes other than simple recombination are
active.  [CHECK THIS]

Now consider the \ion{Fe}{2} transitions.
The bottom left panel of Figure~\ref{fig:fiducial_1d} covers the
majority of the \ion{Fe}{2} UV1 transitions and several are
shown in the velocity plot.  The line
profile for \feiib\ is very similar to those for the \mgiid\ doublet;
one observes strong absorption to negative velocities and strong
emission at $v > 0 \, \mkms$ producing the characteristic P-Cygni profile of
an outflow.  Splitting
the profile at $v = -50 \, \mkms$, we measure an equivalent width
$W_{2600}^{\rm abs} = 1.16$\AA\ in absorption and $W_{2600}^{\rm em} =
-0.94$\AA\ in emission (Table~\ref{tab:fiducial_EW}) for a total
equivalent width of $W_{2600}^{\rm TOT} \approx 0.22$\AA.  
In contrast, the \feiia\ resonance line shows much weaker emission and
a much higher total equivalent width ($W_{2586}^{\rm TOT} = 0.49$\AA)
even though the line has a $2 \times$ lower $f\lambda$ value.
The difference is entirely due to the much lower emission equivalent
width for \feiia\ relative to \feiib.  The \feiia\ emission is reduced
by two factors:
(i) there is an additional downward transition from the
\zconfig$_{7/2}$ level and 
(ii) more importantly, the Einstein A
coefficients of the non-resonant lines are comparable to and even
exceed the Einstein A coefficient of
the resonant transition.  In contrast, 
the \ion{Fe}{2}~2626 transition (associated with \feiib)
has an approximately  $4\times$ smaller coefficient than the
resonance line.  Therefore, the majority of photons absorbed at
$\lambda 2600$ are re-emitted as \feiib\ photons whereas 
the majority of photons absorbed as \feiia\ are re-emitted 
at longer wavelengths (\ion{Fe}{2}$^* \; \lambda 2612$ or $\lambda
2632$).
The total equivalent width, however, of the lines connected to the
\zconfig$_{7/2}$ upper level must still vanish (hotons are conserved
in this fiducial wind model).
Indeed, if were increase $\tau_{2600}$ and especially if we include
gas with $\mvr \approx 0 \mkms$ then the emission at \feiib\ is
significantly suppressed (e.g.\ $\S$~\ref{sec:ISM}).

The preceding discussion emphasize the filling-in of absorption at $v
\lesssim -50 \, \mkms$ and the generation of emission lines at $v \approx
0 \, \mkms$ by photons scattered in the wind.  To study the spatial
extent of this emission, we perfomed 3D calculations with the fiducial
model.  The outpus is a set of surface-brightness maps in a series of
frequency channels, giving a
dataset analagous to integral-field-unit (IFU) observations.  In
Figure~\ref{fig:fiducial_ifu_mgii}, we present the output 
at several velocities relative to the \mgiia\
transition. At $v = -200 \, \mkms$, where the wind has an optical
depth $\tau_{2796} < 1$ (Figure~\ref{fig:fig_fiducial_nvt}),
the source contributes significantly to the observed flux.  
At $v=-100 \, \mkms$, however, the
wind absorbs all photons from the source and the observed emission is
entirely from photons scattered by the wind.  This scattered emission
actually exceeds the source+wind emission at 
$v = -200 \, \mkms$ such
that the absorption profile reaches peak depth at a velocity that is
negatively offset from $\tau_{2796}^{\rm max}$.
The net result of the scattered photons is
a weaker \ion{Mg}{2} profile that is offset from the true optical
depth profile.  Clearly, this complicates estimates for the
speed, covering fraction, and total column density of the wind.  At $v
= 0 \, \mkms$, the wind and source have comparable total flux with the
latter dominating at higher velocities.

Similar characteristics are observed for the \ion{Fe}{2} resonance
transitions (Figure~\ref{fig:fiducial_ifu_feii2600}).
For transitions to fine-structure levels of the \aconfig, the source
is unattenuated but there is also a significant contribution from the
wind (Figure~\ref{fig:fiducial_ifu_feii2612}).

We rexpress these azimuthally symmetric surface-brightness maps in
Figure~\ref{fig:fidicual_cuts} where the emission has been collapsed
along one spatial axis.  For the fiducial model, the majority of
\ion{Mg}{2} emission occurs within the inner few kpc, e.g.\ $50-60\%$ integrating
over $r = \pm 2.5$\,kpc for $v=-100$ to $+100 \, \mkms$.  The fraction
is even more centrally concentrated for the \ion{Fe}{2} transitions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Variations to the Fiducial Model}

In this section, we investigate a series of more complex scenarios
through modifications to the fiducial model.  These include relaxing
the assumption of isotropy, introducing dust, and adding an ISM
component within $r_{\rm inner}$.

\subsection{Anisotropic Winds}
\label{sec:anisotropic}

The fiducial model assumes an
isotropic wind with only radial variations in velocity and density. 
Angular isotropy is obviously an idealized case, but
one that is frequently assumed in studies of galactic-scale outflows
\citep[e.g.][]{steidel+10}.   There are several reasons, however, why
the wind may not be isotropic.  Firstly, galaxies are not spherically
symmetric;  the sources driving the
wind (e.g.\ supernovae, AGN) are very likely to be anisotropically distributed
within the galaxy.  Secondly, the galaxy may be surrounded by a
anisotropic gaseous halo that could preferentially [constrain] the
wind.

With these considerations in mind, we reanalyzed the fiducial model
with the 3D algorithm after setting the density to zero for $2\pi$ steradians.
We then rotated our view of this system from $\theta = 0^\circ$ where
we view the source directly to $\theta = 180^\circ$ where the source
is covered by the anisotropic wind.  The resulting \ion{Mg}{2} and
\ion{Fe}{2} profiles are compared against the fiducial model
(isotropic wind) in Figure~\ref{fig:anisotropic}.

Examining the \mgiid\ doublet, 
there is no attenutation of the source by the wind for $\theta =
0^\circ$
but one does observe significant line emission from photons scattered
off the back side.  These photons, by definition, have $v \gtrsim 0 \mkms$
relative to line-center (a few have $v<0 \mkms$ because
of turbulent motions in the wind). 
When viewed from the opposite direction ($\theta = 180^\circ$), the
absorption lines dominate but there is still significant
line-emission -- at $v \approx 0 \mkms$ and at $v < 0 \mkms$ which fills
in the absorption -- from photons that scatter through the wind.  The
key difference from the isotropic wind is the absence of photons
scatterd to $v > 100 \mkms$ (also note the greater strength of the
\mgiib\ profile at $v \approx -100 \, \mkms$). The asymmetry in emission lines
may therefore provide a robust diagnostic of wind isotropy.
For $\theta = 90^\circ$, the wind covers one half of the source and
one observes half the absorption and also emission at all velocities.

The results are very similar for the \feiid\ resonance lines.  The
\ion{Fe}{2}$^* \; \lambda 2612$ line shows clearly the
separation in velocity for the line-emission.   [Anything else?]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Dust}
\label{sec:dust}

Essentially all astrophysical environments that contain both cool gas
and metals also show signatures of dust depletion and extinction.  This includes the
ISM of star-forming and \ion{H}{1}-selected galaxies
\citep[e.g.][]{ss96,pw01,pcd+07}, strong \ion{Mg}{2} metal-line
absorption systems \citep{york,menard}, and the galactic winds traced
by low-ion transitions \citep{cb58,naI}.  Although the galactic winds
traced specifically by \ion{Mg}{2} and \ion{Fe}{2} transitions have not (yet) been
demonstrated to contain dust, it is reasonable to consider its
effects.

Dust modifies the observed flux in two manners. 
First, it serves as an opacity for the photons
emitted by the source.  This suppresses the flux at all
wavelengths by $\approx \exp(-\mtaud)$, but because we normalize the profiles 
this effect is essentially ignored.  Second, photons that are
scattered by the wind (once or multiple times) into our view will travel a greater
distance and suffer from greater extinction.  A photon that is
trapped for many scatterings may have a high probability of being absorbed
by dust and then exit as an IR photon.  This process is
often the explanation given for the weak (or absent) \lya\ emission
associated with star-forming galaxies \citep[e.g.][]{lya}.
Section~\ref{sec:dust_method} describes our treatment of dust; we only
remind the reader here that the opacity is assumed to scale with the
gas density and that we normalize the extinction by the total optical
depth \taud\ that a photon would experience if it travelled from the
source to infinity without scattering.

In Figure~\ref{fig:dust}, we show the \ion{Mg}{2} and \ion{Fe}{2}
profiles of the fiducial model ($\mtaud = 0$) against a series of
models with $\mtaud > 0$.  For the \ion{Mg}{2} transitions, the
dominant effect is the suppression of line emission at $v \ge 0
\mkms$.  These `red' photons were scattered off the
backside of the wind and must travel a longer path than other
photons.  This differnetial reddening with increasing wavelength (and
velocity relative to line-center) is a natural consequence of dust
extinction and is most evident in the \ion{Fe}{2}$^* \; \lambda 2612$
emission profile.   In terms of the absorption, the profiles are
nearly identical for $\mtaud \le 3$; one requires a very high
extinction to see even a modest effect at $v \approx -100 \, \mkms$.

We conclude that dust is a modest factor for this fiducial model and,
by inference, for models with moderate peak optical depth and
significant velocity gradients with radius (i.e.\ scenarios where the
photons scatter one to a few times before exiting).
At $\mtaud = 10$, however, the source itself is
extinguished by 15\,magnitudes and would never be observed. 
Even $\mtaud = 3$ is larger than typically inferred for the
star-forming galaxies that drive outflows \citep[e.g.][]{dust}.
The dominant effect is a reduction in the flux of the emission lines
with greater extinction at higher velocities relative to line-center.
In these respects, this roughly mimics the behavior of the anistropic
wind described in Section~\ref{sec:anisotropic}. [Is there a
distinguishing factor, e.g. flux of 2612/2600??]

[Do for other density laws, but maybe not show]

\subsection{ISM}
\label{sec:ISM}

The fiducial mdoel does not include gas associated with
the interstellar medium of the galaxy, i.e.\ material at $r \sim
0$\,kpc with $v_{\rm r} \approx 0 \, \mkms$.  We chose this to focus on
properties of the wind.  [word]  The decision was also motivated by
the general absence of significant absorption at $v \approx 0 \mkms$
in galaxies that exhibit outflows
\citep[e.g.][]{wcp+09,steidel+10,rubin+10}.
On the other hand, the stars and \ion{H}{2} regions that comprise our
background sources are very likely embedded within and fueled by gas
of the galactic ISM.  
Consider, then, a modification to the fiducial model that has the
identical wind but also includes an ISM component. 
Specifically, the ISM component has $n_{\rm H} = 1 \cm{-3}$ for
$r_{\rm ISM} \le r < r_{\rm inner}$ with $r_{\rm ISM} = 0.5$\,kpc, 
an average velocity of $\mvr = 0 \mkms$, and a larger turbulent velocity $b_{\rm ISM} = 40 \mkms$.
Figure~\ref{fig:ISM} summarizes the model.
The resultant optical depth profile $\tau_{2796}$ is identical to the
fiducial model for $r > 2$\,kpc but has higher opacity at $r=1-2$\,kpc
and a substantial component at $r = 0.5-1$\,kpc.

In Figure~\ref{fig:ISM_spec}, the solid curves show the \ion{Mg}{2} and
\ion{Fe}{2} profiles for the ISM+wind (black) and fiducial (red)
models. In comparison, the
dotted curve shows the absorption profile for the ISM+wind in the (unphysical) case
where none of the absorbed photons are re-emitted.   Focus first on the
\ion{Mg}{2} doublet.  As expected, the dotted curve shows strong
absorption at $v \approx 0 \mkms$ and blueward.  The full models,
in contrast, show non-zero flux at these velocities and even a
normalized flux exceeding unity at $v \approx 0 \mkms$.  In
fact, the ISM+wind model is nearly identical to the fiducial model;
the only quantitiative difference is that the velocity centroid of
the emission lines are shifted redward by $\approx +100 \, \mkms$.

There are, however, several qualitiative differences 
for the \ion{Fe}{2} transitions between the ISM+wind and fiducial
models.
First,
the \feiia\ transition shows much stronger absorption at $v \approx 0$
to $-100 \mkms$.  In contrast to the \ion{Mg}{2} doublet,
the profile is not filled in by scattered photons. Instead, 
the majority of \feiia\ photons absorbed are re-emitted as
\ion{Fe}{2}~$^* \lambda\lambda 2612, 2632$ photons.  In fact, the
\feiia\ profile with scattering fully included nearly matches the profile without re-emission
(compare the solid and dotted lines); this transition provides a
very good description of the ISM+wind opacity profile.  The other key
difference is the much higher strength ($3-10\times$) of emission from
transitions to the excited states of \aconfig.   There are two reasons
for the higher emission: (1) the absorption is greater in the \feiid\
resonance lines; (2) the high opacity of the ISM component leads to
the increased conversion of resonance photons with $v \approx 0\mkms$
into the non-resonance lines.  This is especially notable for the
\feiib\ transition whose `partner' shows an equivalent width nearly
$10\times$ stronger than for the fiducial model.  The relative
strengths of the \feiib\ and \feiie\ lines therefore provide a direct
diagnostic on the degree to which the resonance lines are trapped.
[Comment:  this is a product of $\tau_{2600}$ and the velocity
gradient]

[3D discussion]

[Fix the bug in the FeII* emission]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Alternate Wind Models}

\subsection{Power-Law Models}
\label{sec:power}

The models investigated in the previous two sections
assumed power-law descriptions for both the density and velocity
expressions of the wind (Equations~\ref{eqn:density},\ref{eqn:vel}).
The power-law exponents were arbitrarily chosen, with minimal physical
motivation.  In this sub-section, we explore the results for a series
of other power-law expressions.

\subsection{Radiation Pressure}

\subsection{The Lyman Break Galaxy Model}
\label{sec:lbg}

The Lyman break galaxies, UV color-selected galaxies at $z \sim 3$,
exhibit cool gas outflows in \ion{Si}{2}, \ion{C}{2},
etc.\ transitions with speeds up to 1000\kms\
\citep[e.g.][]{lowenthal,pettini,steidel+10}.
Researchers have since invoked these winds to explain enrichment of
the intergalactic medium \citep[e.g.][]{aguirre,evan}, the origin of the
damped \lya\ systems \citep{nelson,joop98}, and XXX.  Although the
presence of these outflows were established over a decade ago
\citep{lowenthal,pettini98}, the processes driving the wind remain
unidentified as do [precise] estimates of the mass/energetics of the
outflow.

Recently, \cite[][; hereafter S10]{steidel+10} introduced a model to
explain jointly the average absorption they observed
toward a set of XXX LBGs and the average absorption in cool gas
observed transverse to these galaxies.  
Their wind model is defined by two
expressions: a radial velocity law $\mvr(r)$ and the covering fraction of
optically thick gas $f_c(r)$.  For the latter, S10 envision an
ensemble of small, optically thick clouds $(\tau \gg 1)$ that only
partially cover the galaxy.
For the velocity law, they adopted the following functional
form

\begin{equation}
\mvr = \ltp \frac{A_{\rm LBG}}{1-\alpha} \rtp^{1/2} \ltk r_{\rm
  inner}^{1-\alpha} - r^{1-\alpha} \rtk^{1/2}
\label{eqn:LBG_vlaw}
\end{equation}
with $A_{\rm LBG}$ a constant that sets the terminal speed,
$r_{inner}$ is the inner radius of the wind (taken to be 1\,kpc), and
$\alpha$ describes how steeply the velocity curve rises.  Their
analysis of the LBG absorption profiles implied
a very steeply rising curve with $\alpha \approx 1.3$.
This velocity expression is shown as a dotted line in the top panel of
Figure~\ref{fig:LBG_Sobolev}.

The covering fraction of optically thick cold gas, meanwhile, was assumed to have
the functional form

\begin{equation}
f_c(r) = f_{c,max} \ltp \frac{r}{r_{\rm inner}} \rtp^{-\gamma} \cmma
\label{eqn:covering}
\end{equation}
with $\gamma \approx 0.5$ and $f_{c,max}$ the maximum covering
fraction.  From this expression and the velocity law, one can recover
an absorption profile $I_{\rm LBG}(v) = 1 - f_c(r[v])$, written
explicitly as

\begin{equation}
I_{\rm LBG}(v) = 1 - f_{c,max} \ltk r_{\rm inner}^{1-\alpha} - \ltp
\frac{1-\alpha}{A_{\rm LBG}} \rtp v^2 \rtk^{\gamma}/(\alpha-1)}
\perd
\label{eqn:LBG_I}
\end{equation}
The resulting profile for $f_{c,max} = 0.6$, $\gamma=0.5$,
$\alpha=1.3$, and $A_{\rm LBG} = -192,000 \, \rm km^2 s^{-2} kpc^{-2}$ 
is displayed in the lower panel of
Figure~\ref{fig:LBG_Sobolev}.

In the following, we consider two methods to analyze the LBG wind.
Both approaches assume isotropy and adopt the velocity law given by
Equation~\ref{eqn:LBG_vlaw}.  In one model, we treat the cool gas as a
diffuse medium with unit convering fraction and a radial density
profile determined from the Sobolev approximation.  We then apply 
the methodology used for the other wind models.  In the other model,
we modify our algorithms to more precisely mimic the concept of an
ensemble of optically thick clouds with a partial covering fraction.
by \cite{steidel+10}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Sobolev approximation}
\label{sec:Sobolev}

As demonstrated in Figure~\ref{fig:LBG_Sobolev}, the wind velocity for
the LBG model rises very steeply with increasing radius before
flattening at large radii.  Under these conditions, the Sobolev
approximation relates the optical depth of an expanding medium to the
density and velocity gradient of the gas.  In our case, we 
generate the optical depth prpofile from Equation~\ref{eqn:LBG_I},

\begin{equation}
\tau_{\rm LBG}(v) = -\ln \ltk I_{\rm LBG}(v) \rtk \cmma
\label{eqn:tauLBG}
\end{equation}
and invert the Sobolev approximation to derive the density profile.
Following \cite{lamers+c99}, we have

\begin{equation}
n_{\rm LBG}(r) = \frac{\tau_{\rm LBG}(r) \; d\mvr/dr}{\kappa_\ell
  \lambda}
\label{eqn:Sobolev}
\end{equation}
with $\lambda$ the rest wavelength and $\kappa_\ell \equiv f\pi^2
e^2/m_e c$.

The solid curve in Figure~\ref{fig:LBG_Sobolev}a shows the
resultant density profile for Mg^$+$ assuming that the \mgiia\ line
follows the intensity profile drawn in Figure~\ref{fig:LBG_Sobolev}b. 
This is a relatively extreme density profile.  From the inner radius
of 1\,kpc to 2\,kpc, the density drops by over 2 orders of
magnitude including nearly one order of magnitude over the first
10\,pc.  Beyond 2\,kpc, the density drops even more rapidly, falling
orders of magnitude from 2 to 100\,kpc.

We verified that the density profile shown in
Figure~\ref{fig:LBG_Sobolev}a
reproduces the proper
absorption profile by discretizing the gas into a series of tiny
`clouds' and calculating the integrated absorption profile.  This
calculation is shown as a dotted red curve in
Figure~\ref{fig:LBG_Sobolev}b which is an excellent
match to the desired profile (black curve).
To calculate the optical depth profiles for the other transitions, we
assume $\mnfe = \mnmg/2$ and scale $\tau$ by $f\lambda$.

We generated \ion{Mg}{2} and \ion{Fe}{2} profiles for this wind
using the 1D algorithm with no dust extinction; these are shown as
red curves in Figure~\ref{fig:LBG_spec}.   For this analysis, one
should focus on the \mgiia\ transition.  The dotted line in the Figure
shows the intensity profile for the model when one ignores re-emission
of absorbed photons.  By construction, it follows\footnote{Note that
  one should not make this comparison for the other transitions in
  this Sobolev model because those are scaled down by $f\lambda$ and
  (for \ion{Fe}{2}) the reduced Fe$^+$ abundance.} the profile
described by Equation~\ref{eqn:LBG_I}.   In comparison, the full model
shows much weaker absorption, especially at $v = 0$ to $-300 \mkms$.
In this respect, the LBG-Sobolev model is an
inaccurate description of the observations. The model
also predicts significant emission in the \ion{Mg}{2} lines and several of
the \ion{Fe}{2}$^*$ transitions.   Emission associated with cool gas
has been observed for non-resonant \ion{Si}{2}
transitions in LBGs \citep{cb58,shapley}, but
the \ion{Fe}{2} transitions
modeled here lie in the near-IR and have not yet been investigated.
On the other hand, there have been no 
reported detections of significant line-emission related to resonance
transitions (e.g.\ \ion {Mg}{2}) in LBGs, only for $z \le 1$ star-forming galaxies
\citep{wcp+09,rubin09}.  
[Last line or two]

[Do CIV scattered photons bugger up the P-Cygni analysis of Pettini
et al. ?]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Partial Covering Fraction}
\label{sec:Covering}

In the previous sub-section, we described a Sobolev solution that
reproduces the average absorption profile of LBGs in cool gas
transitions where scattered photons are ignored.  A proper analysis of
this wind model predicts absorption profiles that are qualitatively
different from the ones observed because scattered photons fill-in
absorption and generate significant line emission, similar to the
fiducial wind model ($\S$~\ref{sec:fiducial}).
The LBG-Sobolev model, however, differs from
the one proposed by S10.  They proposed an ensemble of optically
thick clouds with the (parital) covering fraction described by
Equation~\ref{eqn:covering}.   In contrast, the Sobolev model assumes
a diffuse medium with a declining density profile but a unit covering
fraction.  One is left to question whether this difference leads to
the contradictory results of the model. [word]

To more properly model the LBG wind described in S10, we 
performed the following Monte Carlo calculation.  First, we propogated
a photon from the source until its velocity relative to line-center
resonates with the wind (the photon escapes if this never occurs).
The photon then has a probability $P = f_c(r)$ of scattering.  If it
scatters, we track the photon until it comes into
resonance again or escapes the system.  In this model, all of the
resonance transitions are assumed to have identical optical depth in
the absence of re-emission.

The results of the full calculation (absorption plus scattering) are
shown as the black curve in Figure~\ref{fig:LBG_spec}.  The results
are very similar to the Sobolev calculation; scattered photons fill-in
the absorption profiles at $v \approx 0 \mkms$ and yield significant
emission lines at $v \gtrsim 0 \mkms$ in the resonance lines and at
all velocities for the \ion{Fe}{2}$^*$ transitions.  For \feiia,
the majority of absorbed photons have been converted to its
fine-structure counterparts lending to very weak emission at this
transition. The absorption profile very nearly matches the
model without re-emission indicating that this transition well
reproduces the opacity of the wind.  
The equivalent width of \feiia\ even exceeds that for \feiib, an
inversion that if observed would strongly support this model.
[Could SiII or OI do the same?]

Comparing the LBG Sobolev and parital covering models in
Figure~\ref{fig:LBG_spec} (via \mgiia), we note nearly identical
results.  The differencs that occur for the other transitions,
meanwhile, are only because their opacities are scaled downward in the
Sobolev model (as required).  [Do the data prefer one to the other?
And what if we modified the density away from Sobolev?]
In principle, they provide a means to distinguish between the two LBG
models.

[Does most of the scattering occur within the inner few kpc? Explore
this with outflow.cc]

\section{Observables}

Although the focus of this paper is to explore the absorption/emission
profiles for idealized wind models, we ultimately intend to compare 
profiles like these against observations to constrain characteristics
of the wind (e.g.\ raidal extent, energetics).  In this section we
consider a few observables and their dependence on the wind properties.

\subsection{Absorption profiles}

 \begin{enumerate}
   \item Peak optical depth
     \begin{itemize}
       \item Not a robust measure of $\tau_{wind}$
       \item Not even a good relative predictor for $\tau$
       \item Not a robust predictor of $f_c$
     \end{itemize}
   \item Kinematics
     \begin{itemize}
       \item Nearly robust for $v<v_{min}$
       \item But what sets $v_{min}$?
       \item Generally get the full extent of the wind speed
     \end{itemize}
   \item EW
     \begin{itemize}
       \item Primarily set by $v_{min}(\tau \sim 1)$
       \item $f\lambda$ matters, but so do emission channels
     \end{itemize}
   \item Fine-structure
     \begin{itemize}
       \item Distance to the source
     \end{itemize}
 \end{enumerate}

\subsection{Emission profiles}

 \begin{enumerate}
   \item Resonant lines
     \begin{itemize}
       \item Kinematics of backside
       \item Flux + relative
    \end{itemize}
   \item Fine-structure
     \begin{itemize}
       \item Flux $\propto \tau_{wind}$ + relative
       \item Kinematics (anisotropy)
    \end{itemize}
 \end{enumerate}

\section{Discussion}

[Are MgII and FeII* emission strict predictions of outflows?  People
have been so focused on absorption.]

\acknowledgments

J. X. P. and J.M.O. are supported by NASA grant
HST-GO-10878.05-A.  J.X.P and G.W. are partially supported
by an NSF CAREER grant (AST--0548180), and 
by NSF grant AST-0908910.

\bibliographystyle{/u/xavier/paper/Bibli/apj}
\bibliography{/u/xavier/paper/Bibli/allrefs}

\input{Tables/table_atomic.tex}
\input{Tables/tab_fiducial.tex}
\input{Tables/tab_fiducial_ew.tex}
\input{Tables/tab_meas_models.tex}
\input{Tables/tab_meas_plaws.tex}

\begin{figure}
\epsscale{0.8}
\plotone{Figures/energy_levels.ps}
\caption{
Energy level diagrams for the \mgiid\ doublet and the UV1
multiplet of \ion{Fe}{2} transitions   
(based on Figure~7 from \cite{hmt+99}).
Each transition shown is
labelled by its rest wavelength (\AA) and Einstein A-coefficient
(s$^{-1}$). Black updward arrows
indicate the resonance-line transitions connected to the ground
state of each ion.  The 2p$^6$3p configuration of Mg$^+$ is split into
two energy levels that give rise to the \mgiid\ doublet.  
Both the 3d$^6$4s ground state and 3d$^6$4p upper level of Fe$^+$
exhibit fine-structure splitting that gives rise to a series of
electric-dipole transitions. 
The downward (green) arrows show the transitions connected to the
resonance-line transitions (i.e.\ they share the same upper energy
levels).  We also show a pair of levels (\ion{Fe}{2}~$\lambda\lambda
2618,2631$) that arise from higher levels in the \zconfig\
configuration.  These transitions have not yet been observed in
galactic-scale outflows and are not considerd in our analysis.
}
\label{fig:energy}
\end{figure}

\begin{figure}
\epsscale{0.8}
\plotone{Figures/fig_nvtau_vs_r.ps}
\caption{
Density (solid; black), radial velocity (dotted; blue), and
\mgiia\ optical depth profiles (dashed; red) for the fiducial
wind model (see Table~\ref{tab:fiducial} for details).
The density and velocity profiles are simple $r^{-2}$ and $r$
power-laws.  The optical depth profile was calculated by summing
the opacity at small and discrete radial intervals.  One notes that
the wind is optically thick at the inner radius ($r_{\rm inner} =
1$\,kpc) and becomes optically thin at the outer radius ($r_{\rm
  outer} = 20$\,kpc).
Note that the density and velocity curves have been scaled for plotting
convenience.  
}
\label{fig:fiducial_nvt}
\end{figure}

\begin{figure}
\epsscale{0.8}
\plotone{Figures/fig_fiducial_1d.ps}
\caption{
{\it Left} -- (upper) \ion{Mg}{2} profiles for the fiducial wind model
described in Table~\ref{tab:fiducial} and
Figure~\ref{fig:fiducial_nvt}.  The doublet shows the P-Cygni profiles
characteritic of an outflow with significant absorption blueward of
line center (dashed vertical lines) extending to $v = -1000\mkms$
and signficiant emission redward of line center.  Note
that even though the peak optical depth of the \ion{Mg}{2} transitions
is nearly 30 at $v \approx -70 \mkms$, photons scattered off the outflow
fill in the absorption.
(lower) \ion{Fe}{2} absorption and emission profiles for the UV1
multiplet at $\lambda \approx 2600$\AA.  The \feiid\ resonance lines 
show weaker absorption due to the smaller Fe$^+$ number density and
lower $f\lambda$ values.  Each also shows a P-Cygni profile, although
the emission for \feiia\ is significantly weaker than that of the
\feiib\ and \mgiid\ transitions.  This is because a majority of the
absorbed \feiia\ photons are converted into
\ion{Fe}{2}~$\lambda\lambda 2612, 2632$ photons as observed.
[Mention the emission kinematics?]
}
\label{fig:fiducial_1d}
\end{figure}

\begin{figure}
\epsscale{0.8}
\plotone{Figures/fig_noemiss.ps}
\caption{
The black curves show the line profiles (absorption and emission) of
the \ion{Mg}{2} and \ion{Fe}{2} resonance lines for the fiducial wind
model.  These shows the canonical P-Cygni profiles of a source
embedded within an outflow.  Overplotted (red) on each transition is
the predicted absorption profile under the constraint that each
absorbed photon is lost from the system, i.e.\ no scattering or
re-emission.   It is evident that the true profiles have been `filled
in' significantly by light scattered in the wind.  Ignoring this
process, one would model the absorption lines with systematicall lower
optical depth and derive a partial covering. [word]
}
\label{fig:noemiss}
\end{figure}

\begin{figure}
\epsscale{0.8}
\plotone{Figures/fig_fiducial_ifu_mgii.ps}
\caption{
Surface-brightness emission maps around the \mgiia\ transition for the
source+wind complex.  The middel panel shows the 1D spectrum with $v=0
\mkms$ corresponding to $\lambda = 2796.35$\AA\ and the dotted vertical
curves indicating the velocity slices for the emission maps.  The
source has a size $r_{\rm source} = 0.5$\,kpc, traced by a few
pixels at the center.   At $\mvr=-200 \mkms$, the wind has an optical
depth of $\tau_{2796} \approx 1$ such that the source contributes
significantly to the observed flux.  At $\mvr=-100 \mkms$, however, the
wind absorbs all photons from the source and the observed emission is
entirely due to photons scattered by the wind.  Surprisingly, this
emission exceeds that observed at $\mvr = -200\mkms$. At $\mvr \ge 0
\mkms$,  both the source and wind contribute to the observed emission.
}
\label{fig:fiducial_ifu_mgii}
\end{figure}

\begin{figure}
\epsscale{0.8}
\plotone{Figures/fig_fiducial_ifu_feii2600.ps}
\caption{
(a) Surface-brightness emission maps around the \feiib\ transition for the
source+wind complex.  The middle panel shows the 1D spectrum with $v=0
\mkms$ corresponding to $\lambda = 2600.173$\AA\ and the dotted vertical
curves indicating the velocity slices for the emission maps.  
The results are very similar to those observed for the \mgiid\ doublet
(Figure~\ref{fig:fiducial_ifu_mgii}).
}
\label{fig:fiducial_ifu_feii2600}
\end{figure}

\begin{figure*}
\epsscale{0.8}
\plotone{Figures/fig_fiducial_ifu_feii2612.ps}
\caption{
(b) Surface-brightness emission maps around the \ion{Fe}{2}~$\lambda 2612$ transition for the
source+wind complex.  The middle panel shows the 1D spectrum with $v=0
\mkms$ corresponding to $\lambda = 2612.654$\AA\ and the dotted vertical
curves indicating the velocity slices for the emission maps.  
In this case, the source is unattenuated yet scattered photons from
the wind also have a significant contribution. 
}
\label{fig:fiducial_ifu_feii2612}
\end{figure*}

\begin{figure}
\epsscale{0.8}
\plotone{Figures/fig_fiducial_cut.ps}
\caption{
Spatial cuts
}
\label{fig:fiducial_cuts}
\end{figure}

\begin{figure}
\epsscale{0.8}
\plotone{Figures/fig_asymm_spec.ps}
\caption{
Profiles of the \ion{Fe}{2} and \ion{Mg}{2} profiles for the fiducial
case (black lines) compared against an anisotropic wind blowing into
only $2\pi$ steradians viewed from $\theta = 0^\circ$ (source
uncovered) to $\theta = 180^\circ$ (source covered).  One detects
significant emission for all orientations of this wind with the
emission shifting from positive to negative velocities as $\theta$
increases.  
}
\label{fig:anisotropic}
\end{figure}

\begin{figure}
\epsscale{0.8}
\plotone{Figures/fig_dust_spec.ps}
\caption{
Profiles of the \ion{Fe}{2} and \ion{Mg}{2} profiles for the fiducial
model (black) against a series of models that include the effects of
dust extinction.  The primary difference is the suppression of line
emission relative to the continuum (again, normalized to unit value).
A more subtle but important effect is that the redder photons in the
emission lines (corresponding to higher velocity relative to
line-center) suffer greater extinction.  This is most evident in the
\feiic\ emission line and is due to the fact that these photons must
travel further by scattering off the backside of the wind.  Note that
the absorption lines are nearly unmodified until $\mtaud = 10$, a
level of extinction that would preclude observing the source
altogether.
}
\label{fig:dust}
\end{figure}

\begin{figure}
\epsscale{0.8}
\plotone{Figures/fig_ism_diagn.ps}
\caption{
Density (solid; black), radial velocity (dotted; blue), and
\mgiia\ optical depth profiles (dashed; red) for the ISM+wind
model.  The wind is identical to the fiducial model
(Figure~\ref{fig:fiducial_nvt}) but we also include an ISM component
with $n_{\rm H} = 1 \cm{-3}$ that has an average velocity of $\mvr = 0
\mkms$ and a large turbulent velocity $b_{\rm ISM} = 40 \mkms$.
Note that the density and velocity curves have been scaled for plotting
convenience.  
}
\label{fig:ISM}
\end{figure}

\begin{figure}
\epsscale{0.8}
\plotone{Figures/fig_ism_spec.ps}
\caption{
Profiles of the \ion{Fe}{2} and \ion{Mg}{2} profiles for the ISM+wind
model (black) compared against the fiducial wind model (red). 
The dotted line traces the predicted absorption profile in the absence
of any re-emission of absorbed photons.
Regarding the \ion{Mg}{2}~doublet, the primary difference between the
ISM+wind and the fiducial models is the shift of $\approx 100 \mkms$
in the emission lines from $\mvr \approx 0 \mkms$ for the fiducial
model to $\mvr \approx +100\mkms$ for the ISM+wind model. 
[Discuss FeII]
}
\label{fig:ISM_spec}
\end{figure}

\begin{figure}
\epsscale{0.8}
\plotone{Figures/fig_lbg_sobolev.ps}
\caption{
{\it Upper:} The dotted curve shows the velocity law for the LBG model
of \cite{steidel+10}.  Note how rapidly the velocity rises from $r =
1$ to 2\,kpc.  The solid curve shows the Sobolev solution required to
match the average absorption profile of LBG galaxies (lower panel).
The density drops off as $(r-1)^{-0.6}$ 
{\it Lower:} Absorption profiles for the LBG model as defined by
\cite{steidel+10} (black solid curve) and for our Sobolev approximation of
the model (red dotted curve).  
}
\label{fig:LBG_Sobolev}
\end{figure}

\begin{figure}
\epsscale{0.8}
\plotone{Figures/fig_lbg_spec.ps}
\caption{
Profiles of the \ion{Fe}{2} and \ion{Mg}{2} profiles for the fiducial
}
\label{fig:LBG_spec}
\end{figure}

\begin{figure}
\epsscale{0.8}
\plotone{Figures/fig_lbg_cumul.ps}
\caption{
Profiles of the \ion{Fe}{2} and \ion{Mg}{2} profiles for the fiducial
}
\label{fig:LBG_cumul}
\end{figure}

\end{document}